DRAFT \\ From physical principles to classical and quantum \\ Hamiltonian and Lagrangian particle mechanics

### 2017-02-16 2015CompleteDerivation/completeDerivation.tex
Comments from Stephen

The work uses ideas and tools from a broad range of mathematical and scientific subjects: topology \cite{LeeTM}, measure theory \cite{Halmos}, differential geometry (symplectic and Riemannian) \cite{Lee}, group theory and vector spaces \cite{Lang,Young}, statistics \cite{Grimmet}, information theory \cite{Pierce}, Hamiltonian mechanics, Lagrangian mechanics \cite{classical_dynamics}, quantum mechanics \cite{Landau, Weinberg}, thermodynamics \cite{Van Ness}, special relativity \cite{French} and so on.\footnote{The references given provide the reader a list of resources consistent with the development of this work.} Using ideas from all these areas is necessary and valuable, but it does create a number of problems.

@@

In this context, we call the evolution of the system \emph{deterministic} if the state at a given time uniquely identifies states at future times, and \emph{reversible} if it uniquely identifies states at past times. While this is a common enough definition, we need to be clear how this applies to the unstated part. Note that the concept of determinism outlined here is context dependent because the state itself represents only the part of the system that we choose to (or can) describe. In this sense, the unstated part is \emph{always} non-deterministic (and non-reversible) in the sense that the state of the system does not determine its evolution. For example, suppose we define the state as position and momentum of the center of mass of a cannonball. Suppose that the evolution is deterministic on that state. What does it tell us about the cannonball temperature, or about the motion of each of its atoms? Nothing. In this sense, the unstated part is non-deterministic and non-reversible. Could we extend the state and the laws of evolution to account for temperature? Yes, but that would be a different evolution defined on a different state. Does it mean the unstated part is always evolving chaotically? Not at all. The temperature may remain constant throughout the motion of the cannonball. Yet we wouldn't know, since we are not studying it: the evolution is deterministic and/or reversible only as far as the state is concerned. With this in mind, we will restrict ourselves to the cases where the following is valid:

@@

As the term "measurement" has become particularly loaded, let's first characterize what we mean by physical distinguishability in our context.

Consider the motion of a cannonball under inertial and gravitational forces. Light will scatter off of it; as it lands the ground will be deformed and the impact will make the temperature slightly rise. Those external physical processes, which happen no matter what we do, can be used to distinguish the motion of the cannonball as their outcomes are correlated. Therefore we can learn the position by looking at the reflected light, and learn the final kinetic energy by looking at the deformation of the ground. In principle, any external process that has a correlation with the states under study can be used to perform a measurement, and any measurement is based on such a process. That is, for us a measurement is simply a physical process that we can use to distinguish states. Setting up an experiment means choosing a particular process with desired outcomes and forcing the system under study to interact with it one or more times. After that, there is no special role played by the ``observer" in making outcomes come about.

@@

Repeatability is also fundamental. First, to make sure the process is indeed correlated to the states. Second, because ``a single take does not a measurement make". One has to gather enough statistics. Note that the number of takes influences the outcomes: with greater statistics the precision and number of distinguishable cases increases. Therefore the processes, as we defined them, may require repeated interactions with similarly prepared states. They may even be a combination of different kinds of interactions that, taken all together, create a set of distinguishable outcomes. But in the end, however complicated it is, the conceptual model remains the same: each process has a set of possible outcomes, and each possible outcome will be associated with a set of states consistent with that outcome. For example, if the cannonball deformed the ground by this amount then its kinetic energy at impact was within this range; if the electron follows a certain path, then its state is among the ones that have spin $+1/2$.

@@

We claim $\mathcal{S}$ has a topology $\mathsf{T}$. Consider the set of all possible physical outcomes associated with all physical processes. Each possible outcome is associated with a set of states that are compatible with that outcome. Let $\mathsf{T}$ be the set of all sets associated with all physical outcomes. $\mathcal{S} \in \mathsf{T}$ and is associated with the outcome ``the system exists". $\varnothing \in \mathsf{T}$ and is associated with the outcome ``the system doesn't exist". Let $V_1, V_2 \in \mathsf{T}$. Then, by definition, there exists a process $P_1$ that admits $V_1$ as an outcome and a process $P_2$ that admits $V_2$ as an outcome. Consider the process $P$ that combines the outcomes of $P_1$ and $P_2$ with a logical \textsc{and}. This always exists physically as we can prepare the same state multiple times and let it interact with each process separately. $P$ will have a possible outcome $V$ corresponding to the case where $P_1$ gave outcome $V_1$ and $P_2$ gave outcome $V_2$. The states compatible with $V$ must be in both $V_1$ and $V_2$, that is $V = V_1 \cap V_2$. Therefore $\mathsf{T}$ is closed under intersection. Let $V_1, V_2 \in \mathsf{T}$. If they are physically distinguishable, then there exists a physical process $P$ that admits both as outcomes. Given $P$, we can always construct the process $P_0$ that combines $V_1, V_2$ with a logical \textsc{or} into a single outcome $V$, by ``forgetting" which of the two was given. The states compatible with $V$ must be in either $V_1$ or $V_2$, that is $V = V_1 \cup V_2$. Therefore $\mathsf{T}$ is closed under union. $\mathsf{T}$ is a topology by definition.

@@

The generality of this result explains why in physics one always assumes functions to be ``well behaved". As the result was derived from our notion of physical distinguishability, this is not a matter of practical convenience. Suppose we were able to prepare a force field that was zero everywhere in space except at a single point. This gives us a way to tag a specific point. But it also allows us to create an outcome compatible with only a single point. A finite precision measurement of the force provides us infinite precision of space, which we ruled out.\footnote{This assumes that we are able to position the probe perfectly, which we could do if we were able to manipulate forces at that precision.} This is what the math is telling us, that if we claim that outcomes can't distinguish isolated points (i.e. standard topology on $\mathbb{R}^n$) then neither can maps. It is physical consistency that limits us to continuous functions.

@@

Evolved state variables are useful to write equations of motions, study how physical quantities change, form a physical picture of what happens. Transported state variables are useful to write invariants, study state space trajectories, form a geometric picture for the state space.\footnote{One should not confuse evolved/transported state variable with active/passive transformations or with Schroedinger/Heisenberg pictures. In those cases, the choice is between changing the state or the coordinates/observables. In our case, the state is always changing. The choice is between tracking the change with different values of the same state variable or different state variables that give the same value.}

@@

On a state space with countable elements, using a single state variable, we have $\Lambda_U (\mathcal{c}) = \sum \limits_{q=a}^b \rho_\mathcal{c}(q)$. We would expect the expression to become $\Lambda_U (\mathcal{c}) = \int_a^b \rho_\mathcal{c} (q) dq$ for states identified by a continuous state variable. By changing state variable, though, we see that this does not work in general: $\rho_\mathcal{c}(\hat{q})= \rho_\mathcal{c}(q) dq/d\hat{q}$. This makes $\rho_\mathcal{c}$ a function not just of the state $\mathcal{s}$, but also of the state variables we are using. This is inconsistent with our previous definition. Moreover if the transformation is not differentiable, the density is not even well-defined. This tells us that $\mathcal{S}$ cannot be any manifold: it has to be one that allows state-variable-independent densities.

@@

    Even if we are able to track parts, the assumption requires them to be infinitesimal, the limit of a process of infinite recursive subdivision. The best we can do experimentally is to confirm that the assumption holds up to the smallest precision available. Therefore even in the best of cases it cannot be considered an experimentally validated assumption but a reasonable default assumption (e.g. "it worked so far").

@@

We have a right triangle-like relationship between the plane where $\omega$ is defined and its projections on the planes defined by each d.o.f., similar to the multiple d.o.f.:
\begin{align*}
\textrm{multiple d.o.f.} \;\;\; &dq^1 \wedge dp_1 + dq^2 \wedge dp_2 = \omega \\
\textrm{temporal d.o.f.} \;\;\; &dt \wedge dE + \omega = dq \wedge dp
\end{align*}
But in the previous case, the right angle was between the two independent d.o.f.. In this case, the right angle is between $\omega$ and the plane of constant $(q, p)$ where $dt \wedge dE$ is defined. We rewrite it as $dq \wedge dp - dt \wedge dE = \omega$. This corresponds to the Minkowski product across d.o.f. and the vector product within. In terms of unified state variables $\xi^a\equiv \{t, q^i, E, p_i\}$ we have:

@@

	We claim $d_s \rho_{\bar{\mathcal{c}}} = \{ \rho_{\bar{\mathcal{c}}}, \mathcal{H} \} = 0$. $d_s \rho_{\bar{\mathcal{c}}}= \partial_a \rho_{\bar{\mathcal{c}}} d_s \xi^a = \partial_a \rho_{\bar{\mathcal{c}}} \omega^{ab} \partial_b \mathcal{H} = \{ \rho_{\bar{\mathcal{c}}}, \mathcal{H} \}$. Let $\gamma : \mathbb{R} \rightarrow T^*\mathcal{M}$ be a parametrization for a particle evolution $\bar{\mathcal{s}}$. Then $\rho_{\bar{\mathcal{c}}}(\lambda(s)) = \rho_{\bar{\mathcal{c}}}(\lambda(s + \Delta s))$ as the material does not change over the particle evolution. $d_s \rho_{\bar{\mathcal{c}}} = 0$.

\end{justification}

@@

	The first natural question to ask is: why does state evolution of a particle of a classical material always describe a trajectory in space? It's because spatial extent is the one way in which all materials can be decomposed. As we divide materials into smaller pieces, physical location is one state variable we can assign to the particles of all materials. If we assume kinematic equivalence, the kinematic variables will be enough to identify a particle state. In such a case, we will call the material \emph{minimal} to highlight the idea that the state is described by the smallest possible number of d.o.f..

	When the system is not minimal, the state of the particles will be described by some additional independent state variables. The overall dynamics will be the spatial d.o.f. coupled with the internal d.o.f.. Therefore by studying a minimal material, we are also studying a component of the dynamics that is present in all materials. In other words: any material is an extension of a minimal material.

@@

Since both $dx^\alpha$ and $du^\alpha$ are four vector components, $g$ is a rank two tensor on $\mathcal{M}$. To each direction in space-time $dx^\alpha$ is associated a direction in the tangent space of velocities $du^\alpha$ with which it forms a d.o.f.. Across d.o.f., that is $\alpha \neq \beta$, we have $\omega(dq^\alpha, dp_\beta) = 0 =\omega(dq^\beta, dp_\alpha)$. Therefore if $dx^\alpha$ and $du^\beta$ do not form a d.o.f., $g(dx^\alpha, du^\beta) = 0 = g(dx^\beta, du^\alpha)$: $g$ is reflexive. We also know that $g$ cannot be alternating, that is $g(dx^\alpha, du^\alpha) \neq 0$: a choice of units for $x^\alpha$ is a choice of units for $u^\alpha$ and therefore they cannot belong to different d.o.f. This means $g$ is a symmetric non-degenerate tensor. We recognize it as the metric tensor and therefore $\mathcal{M}$ is a pseudo-Riemannian manifold.

Since $g$ is symmetric, it can be diagonalized: $\omega(dq^\alpha, dp_\beta)= dq^i \omega_{q^ip_i} dp_i + dt \omega_{tE} dE \propto g(dx^\alpha, du^\beta) = dx^i g_{ii} du^i + dt g_{00} du^0$. As $\omega_{q^ip_i} > 0$ we must have $g_{ii} > 0$. As $\omega_{tE} < 0$ we must have $g_{00} < 0$. We can further change the units along the spatial directions so that $g_{ii}=1$: if we use the same units for the spatial coordinates, each direction will contribute the same number of possibilities. Since time must use a different unit, we set $g_{00}=-c^2$ where $c^2$ is the constant that allows one to convert the possibility count over the temporal d.o.f. into the possibility count of a spatial d.o.f.. Therefore, locally $g(dx^i, dx^i) = (dx^i)^2 - c^2 dt^2$ is Minkowskian.

@@

\begin{justification}
	We claim $(\mathcal{M}, g)$ is a pseudo-Riemannian manifold where $g$ is a metric tensor induced by $\omega$ under kinematic equivalence. Let $(dx^\alpha, du^\alpha)$ be an infinitesimal range of possibilities for the kinematic variables of the particles of a minimal classical material. There exists a map $g : T\mathcal{M} \times T\mathcal{M} \rightarrow \mathbb{R}$ such that $g(dx^\alpha, du^\alpha) \propto \omega(dq^\alpha, dp_\alpha)$. The map $g$ is linear, since a linear combination of infinitesimal ranges leads to the linear combination of possibilities. $g$ is a rank two tensor. Suppose $g(dx^\alpha, du^\beta) = 0$. $g(dx^\alpha, du^\beta) \propto \omega(dq^\alpha, dp_\beta) = 0 = \omega(dq^\beta, dp_\alpha) \propto g(dx^\beta, du^\alpha) = 0$. $g$ is reflexive. Position and velocity along the same direction use related units. They cannot form independent d.o.f.. $g(dx^\alpha, du^\alpha) \neq 0$. $g$ is not alternating. $g$ is symmetric. $g$ is a symmetric rank two tensor. $(\mathcal{M}, g)$ is a pseudo-Riemannian manifold.

@@

The metric tensor $g$ and the constant $c$ found here are indeed the ones used in the context of special relativity. Yet, they have acquired a different meaning. $g$ does not measure space-time distances: it counts possibilities of a d.o.f., like $\omega$. Directions in space are orthogonal not because the angle between them is $90^\circ$, but because they identify independent d.o.f.. That is, the choice of units along one direction is independent from the choice of units along orthogonal directions. Space and time, instead, are not independent d.o.f. and the geometry between them is different. In other words: $g$ inherits its signature from $\omega$. The geometry of space-time is the geometry of the ranges for possible states.

@@

This also allows us to understand heuristically why the speed of deterministic and reversible motion cannot be greater than $c$. Consider the trajectory of a point particle. Over an interval $dt$ it will travel $dx$ in space. Since the motion is deterministic and reversible, and therefore continuous, it will travel through all the states defined in the interval $dx$. It will also travel through every moment in time, which corresponds to $cdt$ possibilities. As it moves, time will always change but not necessarily space. The number of states traveled in space cannot exceed the number of states traveled in time. Therefore we have:
\begin{align*}
\frac{|n_x(dx)|}{|n_t(dt)|} &=  \frac{|dx|}{|c dt|} \leq 1 \\
\frac{|dx|}{|dt|} & \leq c \\
\end{align*}
Note, though, that this constraint applies only to deterministic and reversible motion. It was under that premise that we were able to show that the trajectories are continuous, that we could define $\omega$ and therefore $g$. Non-local quantum effects, such as the ones featured in EPR\cite{EPR} experiments, are not examples of deterministic and reversible evolution. Therefore they do not have anything to violate: the maximum speed does not apply to their ``motion".\footnote{The very notion of velocity becomes ill defined in such cases precisely because we do not have a unique well-defined evolution.} On the other hand, only deterministic and reversible motion can carry information and can establish before/after relationships, therefore everything remains consistent.

@@

	We claim the Hamiltonian is $\mathcal{H}=\frac{1}{2m}((p_\alpha-A_\alpha)g^{\alpha\beta}(p_\beta-A_\beta) + m^2 c^2)$.  $\frac{1}{m}g^{\alpha\beta}(p_\beta-A_\beta) = u^\alpha = d_s x^\alpha = d_s q^\alpha = \partial_{p_\alpha} \mathcal{H}$. After integration $\mathcal{H}=\frac{1}{2m}(p_\alpha-A_\alpha)g^{\alpha\beta}(p_\beta-A_\beta)+V(x)$ where $V$ is an arbitrary function. Let $s$ be the natural parametrization such that $u^\alpha u_\alpha = - c^2$. $V= \frac{1}{2}m c^2$ as $\mathcal{H}=0$ by \ref{prop:form_of_invariant_hamiltonian}. $\mathcal{H}=\frac{1}{2m}((p_\alpha-A_\alpha)g^{\alpha\beta}(p_\beta-A_\beta) + m^2 c^2)$.

\end{justification}

@@

We should also ask whether any choice of $A$ actually represents a different case leading to different trajectories. The simplest and more insightful way to study this problem is to use the kinematic quantities $x^\alpha$ and $mu_\alpha$ as state variables. They are not going to be conjugates of each other but they do capture the entire physical description. Any change in $A$ that does not affect the relationship between $x^\alpha$ and $mu_\alpha$ does not change the physics. Expressing the Hamiltonian and the Poisson brackets in terms of the kinematic variables we have:
\begin{align*}
\mathcal{H}&=\frac{1}{2}mu_\alpha u^\alpha + \frac{1}{2} m c^2 \\
\{x^\alpha, x^\beta\} &= 0 \\
\{x^\alpha, mu_\beta\} &= \delta^\alpha_\beta \\
\{mu_\alpha, mu_\beta\} &= \{p_\alpha, - A_\beta\} + \{-A_\alpha, p_\beta\} \\
&= \partial_\alpha A_\beta - \partial_\beta A_\alpha = F_{\alpha\beta}
\end{align*}
Note how the constraint $\mathcal{H}=0$ simply sets the rest kinetic energy. Note how the Poisson bracket between kinetic momentum is the force tensor. The weaker the force, the more $(x^i, mu_i)$ become independent d.o.f. The force tensor, then, tells how much $(x^i, mu_i)$ fail to be independent d.o.f.. With these expressions, the Poisson bracket between any function of position and kinetic momentum can now be calculated. For example:
\begin{align*}
d_s x^\alpha &= \{x^\alpha, \mathcal{H}\} = u^\alpha \\
d_s (mu^\alpha) &= \{mu^\alpha, \mathcal{H}\} = F^{\alpha\beta} u_\beta
\end{align*}
where we recognize the covariant form of the Lorentz force.

@@

The other feature is particle-wave duality. If the spread of the distribution is narrow enough for the problem at hand, the motion can be described by just the position and momentum of the center of mass, making it behave like a classical particle. If the spread is not negligible (e.g. the object encounters obstacles of a smaller scale) then we have to use the full distribution. Therefore, even in this classical approximation, the electron would have this dual nature depending on the problem.

@@

The generalization to multiple d.o.f. is outside of the scope of this work. Note, though, that this result is related to the Gromov non-squeezing theorem and the symplectic camel, which are already generalized to multiple d.o.f.. It is the link with information theory that gives us further physical insight and is of particular interest given the strong connection between thermodynamics and Shannon's entropy.

>>>>

The work uses ideas and tools from a broad range of mathematical and scientific subjects: topology~\cite{LeeTM}, measure theory~\cite{Halmos}, differential geometry (symplectic and Riemannian)~\cite{Lee}, group theory and vector spaces~\cite{Lang,Young}, statistics~\cite{Grimmet}, information theory~\cite{Pierce}, Hamiltonian mechanics, Lagrangian mechanics~\cite{classical_dynamics}, quantum mechanics~\cite{Landau, Weinberg}, thermodynamics~\cite{Van Ness}, special relativity \cite{French} and so on.\footnote{The references given provide the reader a list of resources consistent with the development of this work.} Using ideas from all these areas is necessary and valuable, but it does create a number of problems.

@@

In this context, we call the evolution of the system \emph{deterministic} if the state at a given time uniquely identifies states at future times, and \emph{reversible} if it uniquely identifies states at past times. While this is a common enough definition, we need to be clear how this applies to the unstated part. Note that the concept of determinism outlined here is context dependent because the state itself represents only the part of the system that we choose to (or can) describe. In this sense, the unstated part is \emph{always} non-deterministic (and non-reversible) in the sense that the state of the system does not determine its evolution. For example, suppose we define the state as position and momentum of the center of mass of a cannonball. Suppose that the evolution is deterministic on that state. What does it tell us about the cannonball temperature, or about the motion of each of its atoms? Nothing. In this sense, the unstated part is non-deterministic and non-reversible. Could we extend the state and the laws of evolution to account for temperature? Yes, but that would be a different evolution defined on a different state. Does it mean the unstated part is always evolving randomly? Not at all. The temperature may remain constant throughout the motion of the cannonball. Yet we wouldn't know, since we are not studying it: the evolution is deterministic and/or reversible only as far as the state is concerned. With this in mind, we will restrict ourselves to the cases where the following is valid:

@@

As the term ``measurement" has become particularly loaded, let's first characterize what we mean by physical distinguishability in our context.

Consider the motion of a cannonball under inertial and gravitational forces. Light will scatter off of it; as it lands, the ground will be deformed and the impact will make the temperature rise slightly. Those external physical processes, which happen no matter what we do, can be used to distinguish the motion of the cannonball as their outcomes are correlated. Therefore we can learn the position by looking at the reflected light, and learn the final kinetic energy by looking at the deformation of the ground. In principle, any external process that has a correlation with the states under study can be used to perform a measurement, and any measurement is based on such a process. That is, for us a measurement is simply a physical process that we can use to distinguish states. Setting up an experiment means choosing a particular process with desired outcomes and forcing the system under study to interact with it one or more times. After that, there is no special role played by the ``observer" in making outcomes come about.

@@

Repeatability is also fundamental. First, to make sure the process is indeed correlated to the states. Second, because ``a single take does not a measurement make." One has to gather enough statistics. Note that the number of takes influences the outcomes: with greater statistics the precision and number of distinguishable cases increases. Therefore the processes, as we defined them, may require repeated interactions with similarly prepared states. They may even be a combination of different kinds of interactions that, taken all together, create a set of distinguishable outcomes. But in the end, however complicated it is, the conceptual model remains the same: each process has a set of possible outcomes, and each possible outcome will be associated with a set of states consistent with that outcome. For example, if the cannonball deformed the ground by this amount then its kinetic energy at impact was within this range; if the electron follows a certain path, then its state is among the ones that have spin $+1/2$.

@@

We claim $\mathcal{S}$ has a topology $\mathsf{T}$. Consider the set of all possible physical outcomes associated with all physical processes. Each possible outcome is associated with a set of states that are compatible with that outcome. Let $\mathsf{T}$ be the set of all sets associated with all physical outcomes. $\mathcal{S} \in \mathsf{T}$ and is associated with the outcome ``the system exists." $\varnothing \in \mathsf{T}$ and is associated with the outcome ``the system doesn't exist." Let $V_1, V_2 \in \mathsf{T}$. Then, by definition, there exists a process $P_1$ that admits $V_1$ as an outcome and a process $P_2$ that admits $V_2$ as an outcome. Consider the process $P$ that combines the outcomes of $P_1$ and $P_2$ with a logical \textsc{and}. This always exists physically as we can prepare the same state multiple times and let it interact with each process separately. $P$ will have a possible outcome $V$ corresponding to the case where $P_1$ gave outcome $V_1$ and $P_2$ gave outcome $V_2$. The states compatible with $V$ must be in both $V_1$ and $V_2$, that is $V = V_1 \cap V_2$. Therefore $\mathsf{T}$ is closed under intersection. Let $V_1, V_2 \in \mathsf{T}$. If they are physically distinguishable, then there exists a physical process $P$ that admits both as outcomes. Given $P$, we can always construct the process $P_0$ that combines $V_1, V_2$ with a logical \textsc{or} into a single outcome $V$, by ``forgetting" which of the two was given. The states compatible with $V$ must be in either $V_1$ or $V_2$, that is $V = V_1 \cup V_2$. Therefore $\mathsf{T}$ is closed under union. $\mathsf{T}$ is a topology by definition.

@@

The generality of this result explains why in physics one always assumes functions to be ``well behaved." As the result was derived from our notion of physical distinguishability, this is not a matter of practical convenience. Suppose we were able to prepare a force field that was zero everywhere in space except at a single point. This gives us a way to tag a specific point. But it also allows us to create an outcome compatible with only a single point. A finite precision measurement of the force provides us infinite precision of space, which we ruled out.\footnote{This assumes that we are able to position the probe perfectly, which we could do if we were able to manipulate forces at that precision.} This is what the math is telling us, that if we claim that outcomes can't distinguish isolated points (i.e. standard topology on $\mathbb{R}^n$) then neither can maps. It is physical consistency that limits us to continuous functions.

@@

Evolved state variables are useful to write equations of motion, study how physical quantities change, form a physical picture of what happens. Transported state variables are useful to write invariants, study state space trajectories, form a geometric picture for the state space.\footnote{One should not confuse evolved/transported state variable with active/passive transformations or with Schroedinger/Heisenberg pictures. In those cases, the choice is between changing the state or the coordinates/observables. In our case, the state is always changing. The choice is between tracking the change with different values of the same state variable or different state variables that give the same value.}

@@

On a state space with countable elements, using a single state variable, we have $\Lambda_U (\mathcal{c}) = \sum \limits_{q=a}^b \rho_\mathcal{c}(q)$. We would expect the expression to become $\Lambda_U (\mathcal{c}) = \int_a^b \rho_\mathcal{c} (q) dq$ for states identified by a continuous state variable. By changing state variable, though, we see that this does not work in general: $\rho_\mathcal{c}(\hat{q})= \rho_\mathcal{c}(q) dq/d\hat{q}$. For example, if we changed units from $g/m$ to $g/Km$, the value of the density would increase by a factor of $1000$. This makes $\rho_\mathcal{c}$ a function not just of the state $\mathcal{s}$, but also of the state variables we are using. This is inconsistent with our previous definition. Moreover if the transformation is not differentiable, the density is not even well-defined. This tells us that $\mathcal{S}$ cannot be any manifold: it has to be one that allows state-variable-independent densities.

@@

    Even if we are able to track parts, the assumption requires them to be infinitesimal, the limit of a process of infinite recursive subdivision. The best we can do experimentally is to confirm that the assumption holds up to the smallest precision available. Therefore even in the best of cases it cannot be considered an experimentally validated assumption but a reasonable default assumption (e.g. ``it worked so far").

@@

We have a right triangle-like relationship between the plane where $\omega$ is defined and its projections on the planes defined by each d.o.f., similar to the multiple d.o.f.:
\begin{align*}
\textrm{multiple d.o.f.} \;\;\; &dq^1 \wedge dp_1 + dq^2 \wedge dp_2 = \omega \\
\textrm{temporal d.o.f.} \;\;\; &dt \wedge dE + \omega = dq \wedge dp
\end{align*}
But in the previous case, the right angle was between the two independent d.o.f. In this case, the right angle is between $\omega$ and the plane of constant $(q, p)$ where $dt \wedge dE$ is defined. We rewrite it as $dq \wedge dp - dt \wedge dE = \omega$. This corresponds to the Minkowski product across d.o.f. and the vector product within. In terms of unified state variables $\xi^a\equiv \{t, q^i, E, p_i\}$ we have:

@@

	We claim $d_s \rho_{\bar{\mathcal{c}}} = \{ \rho_{\bar{\mathcal{c}}}, \mathcal{H} \} = 0$. $d_s \rho_{\bar{\mathcal{c}}}= \partial_a \rho_{\bar{\mathcal{c}}} d_s \xi^a = \partial_a \rho_{\bar{\mathcal{c}}} \omega^{ab} \partial_b \mathcal{H} = \{ \rho_{\bar{\mathcal{c}}}, \mathcal{H} \}$. Let $\gamma : \mathbb{R} \rightarrow T^*\mathcal{M}$ be a parametrization for a particle evolution $\bar{\mathcal{s}}$. Then $\rho_{\bar{\mathcal{c}}}(\lambda(s)) = \rho_{\bar{\mathcal{c}}}(\lambda(s + \Delta s))$ as the material does not change over the particle evolution. $d_s \rho_{\bar{\mathcal{c}}} = 0$.
\end{justification}

@@

	The first natural question to ask is: why does state evolution of a particle of a classical material always describe a trajectory in space? It's because spatial extent is the one way in which all materials can be decomposed. As we divide materials into smaller pieces, physical location is one state variable we can assign to the particles of all materials. If we assume kinematic equivalence, the kinematic variables will be enough to identify a particle state. In such a case, we will call the material \emph{minimal} to highlight the idea that the state is described by the smallest possible number of d.o.f.

	When the system is not minimal, the state of the particles will be described by some additional independent state variables. The overall dynamics will be the spatial d.o.f. coupled with the internal d.o.f. Therefore by studying a minimal material, we are also studying a component of the dynamics that is present in all materials. In other words: any material is an extension of a minimal material.

@@

Since both $dx^\alpha$ and $du^\alpha$ are four vector components, $g$ is a rank two tensor on $\mathcal{M}$. To each direction in space-time $dx^\alpha$ is associated a direction in the tangent space of velocities $du^\alpha$ with which it forms a d.o.f. Across d.o.f., that is $\alpha \neq \beta$, we have $\omega(dq^\alpha, dp_\beta) = 0 =\omega(dq^\beta, dp_\alpha)$. Therefore if $dx^\alpha$ and $du^\beta$ do not form a d.o.f., $g(dx^\alpha, du^\beta) = 0 = g(dx^\beta, du^\alpha)$: $g$ is reflexive. We also know that $g$ cannot be alternating, that is $g(dx^\alpha, du^\alpha) \neq 0$: a choice of units for $x^\alpha$ is a choice of units for $u^\alpha$ and therefore they cannot belong to different d.o.f. This means $g$ is a symmetric non-degenerate tensor. We recognize it as the metric tensor and therefore $\mathcal{M}$ is a pseudo-Riemannian manifold.

Since $g$ is symmetric, it can be diagonalized: $\omega(dq^\alpha, dp_\beta)= dq^i \omega_{q^ip_i} dp_i + dt \omega_{tE} dE \propto g(dx^\alpha, du^\beta) = dx^i g_{ii} du^i + dt g_{00} du^0$. As $\omega_{q^ip_i} > 0$ we must have $g_{ii} > 0$. As $\omega_{tE} < 0$ we must have $g_{00} < 0$. We can further change the units along the spatial directions so that $g_{ii}=1$: if we use the same units for the spatial coordinates, each direction will contribute the same number of possibilities. Since time must use a different unit, we set $g_{00}=-c^2$ where $c^2$ is the constant that allows one to convert the possibility count over the temporal d.o.f. into the possibility count of a spatial d.o.f. Therefore, locally $g(dx^i, dx^i) = (dx^i)^2 - c^2 dt^2$ is Minkowskian.

@@

\begin{justification}
	We claim $(\mathcal{M}, g)$ is a pseudo-Riemannian manifold where $g$ is a metric tensor induced by $\omega$ under kinematic equivalence. Let $(dx^\alpha, du^\alpha)$ be an infinitesimal range of possibilities for the kinematic variables of the particles of a minimal classical material. There exists a map $g : T\mathcal{M} \times T\mathcal{M} \rightarrow \mathbb{R}$ such that $g(dx^\alpha, du^\alpha) \propto \omega(dq^\alpha, dp_\alpha)$. The map $g$ is linear, since a linear combination of infinitesimal ranges leads to the linear combination of possibilities. $g$ is a rank two tensor. Suppose $g(dx^\alpha, du^\beta) = 0$. $g(dx^\alpha, du^\beta) \propto \omega(dq^\alpha, dp_\beta) = 0 = \omega(dq^\beta, dp_\alpha) \propto g(dx^\beta, du^\alpha) = 0$. $g$ is reflexive. Position and velocity along the same direction use related units. They cannot form independent d.o.f. $g(dx^\alpha, du^\alpha) \neq 0$. $g$ is not alternating. $g$ is symmetric. $g$ is a symmetric rank two tensor. $(\mathcal{M}, g)$ is a pseudo-Riemannian manifold.

@@

The metric tensor $g$ and the constant $c$ found here are indeed the ones used in the context of special relativity. Yet, they have acquired a different meaning. $g$ does not measure space-time distances: it counts possibilities of a d.o.f., like $\omega$. Directions in space are orthogonal not because the angle between them is $90^\circ$, but because they identify independent d.o.f. That is, the choice of units along one direction is independent from the choice of units along orthogonal directions. Space and time, instead, are not independent d.o.f. and the geometry between them is different. In other words: $g$ inherits its signature from $\omega$. The geometry of space-time is the geometry of the ranges for possible states.

@@

This also allows us to understand heuristically why the speed of deterministic and reversible motion cannot be greater than $c$. Consider the trajectory of a point particle. Over an interval $dt$ it will travel $dx$ in space. Since the motion is deterministic and reversible, and therefore continuous, it will travel through all the states defined in the interval $dx$. It will also travel through every moment in time, which corresponds to $cdt$ possibilities. As it moves, time will always change but not necessarily space. The number of states traveled in space cannot exceed the number of states traveled in time. Therefore we have:
\begin{align*}
\frac{|n_x(dx)|}{|n_t(dt)|} &=  \frac{|dx|}{|c dt|} \leq 1 \\
\frac{|dx|}{|dt|} & \leq c \\
\end{align*}
Note, though, that this constraint applies only to deterministic and reversible motion. It was under that premise that we were able to show that the trajectories are continuous, that we could define $\omega$ and therefore $g$. Non-local quantum effects, such as the ones featured in EPR\cite{EPR} experiments, are not examples of deterministic and reversible evolution. Therefore they do not have anything to violate: the maximum speed does not apply to their ``motion."\footnote{The very notion of velocity becomes ill defined in such cases precisely because we do not have a unique well-defined evolution.} On the other hand, only deterministic and reversible motion can carry information and can establish before/after relationships, therefore everything remains consistent.

@@

	We claim the Hamiltonian is $\mathcal{H}=\frac{1}{2m}((p_\alpha-A_\alpha)g^{\alpha\beta}(p_\beta-A_\beta) + m^2 c^2)$.  $\frac{1}{m}g^{\alpha\beta}(p_\beta-A_\beta) = u^\alpha = d_s x^\alpha = d_s q^\alpha = \partial_{p_\alpha} \mathcal{H}$. After integration $\mathcal{H}=\frac{1}{2m}(p_\alpha-A_\alpha)g^{\alpha\beta}(p_\beta-A_\beta)+V(x)$ where $V$ is an arbitrary function. Let $s$ be the natural parametrization such that $u^\alpha u_\alpha = - c^2$. $V= \frac{1}{2}m c^2$ as $\mathcal{H}=0$ by \ref{prop:form_of_invariant_hamiltonian}. $\mathcal{H}=\frac{1}{2m}((p_\alpha-A_\alpha)g^{\alpha\beta}(p_\beta-A_\beta) + m^2 c^2)$.
\end{justification}

@@

We should also ask whether any choice of $A$ actually represents a different case leading to different trajectories. The simplest and more insightful way to study this problem is to use the kinematic quantities $x^\alpha$ and $mu_\alpha$ as state variables. They are not going to be conjugates of each other but they do capture the entire physical description. Any change in $A$ that does not affect the relationship between $x^\alpha$ and $mu_\alpha$ does not change the physics. Expressing the Hamiltonian and the Poisson brackets in terms of the kinematic variables we have:
\begin{align*}
\mathcal{H}&=\frac{1}{2}mu_\alpha u^\alpha + \frac{1}{2} m c^2 \\
\{x^\alpha, x^\beta\} &= 0 \\
\{x^\alpha, mu_\beta\} &= \delta^\alpha_\beta \\
\{mu_\alpha, mu_\beta\} &= \{p_\alpha, - A_\beta\} + \{-A_\alpha, p_\beta\} \\
&= \partial_\alpha A_\beta - \partial_\beta A_\alpha = F_{\alpha\beta}
\end{align*}
Note how the constraint $\mathcal{H}=0$ simply sets the rest kinetic energy. Note how the Poisson bracket between kinetic momentum is the force tensor. The weaker the force, the more $(x^i, mu_i)$ become independent d.o.f. The force tensor, then, tells how much $(x^i, mu_i)$ fail to be independent d.o.f. With these expressions, the Poisson bracket between any function of position and kinetic momentum can now be calculated. For example:
\begin{align*}
d_s x^\alpha &= \{x^\alpha, \mathcal{H}\} = u^\alpha \\
d_s (mu^\alpha) &= \{mu^\alpha, \mathcal{H}\} = F^{\alpha\beta} u_\beta
\end{align*}
where we recognize the covariant form of the Lorentz force.

@@

The other feature is particle-wave duality. If the spread of the distribution is narrow enough for the problem at hand, the motion can be described by just the position and momentum of the center of mass, making it behave like a classical particle. If the spread is not negligible (e.g. the object encounters obstacles or slits of a smaller scale) then we have to use the full distribution. Therefore, even in this classical approximation, the electron would have this dual nature depending on the problem.

@@

The generalization to multiple d.o.f. is outside of the scope of this work. Note, though, that this result is related to the Gromov non-squeezing theorem and the symplectic camel, which are already generalized to multiple d.o.f. It is the link with information theory that gives us further physical insight and is of particular interest given the strong connection between thermodynamics and Shannon's entropy.
